---
title: "fars_functions"
author: "R. Townsend"
date: "2021-09-11"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fars_functions_details}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(PracticePkg)
library(readr)
library(dplyr)
```
The functions in the fars_function package will assist you in querying the **US National Highway Traffic Safety Administration's Fatality Analysis Reporting System ** (*FARS*) for annual reported accidents, summarize them by month and year, and plot teh accident sites for a selected state and year.

Individual files can be read in by using 

```{r, eval=FALSE}
fars_read( year = "accident_2013.csv.bz2")
```
which is a bz2 compressed file from the FARS database for 2013.  


This is a bit unwieldy for querying multiple years, so the function **make_file** is used to create appropriate file name(s) based on the requested year(s) for the more advanced functions:

```{r, eval=FALSE}
make_filename(2013:2015)
```

**fars_read_years** queries multiple *years* using fars_read_years returns a tibble with the month and year of each reported accident.
```{r, eval=FALSE}
fars_read_years( years = 2013:2015)
```


Summarization of annual reported accident counts into one month-by-year tibble is performed by the **fars_summarize_years** function. 

```{r, eval=FALSE}
fars_summarize_years( years = 2013:2015)
```

For a plot of annual accidents reported for a selected state, use the **fars_map_state** function.
```{r, eval=FALSE}
fars_map_state( state.num = 16, year = 2013)
```


